
```{r}
# Set parameters
theta_true <- 3
n <- 1000  # sample size
N_attempts <- 1000  # number of simulation attempts

# Function to calculate MLE for Poisson distribution
mle_poisson <- function(data) {
  mean(data)
}

# Initialize vector for results
results <- numeric(N_attempts)

# Simulate and calculate MLE multiple times
for (i in 1:N_attempts) {
  simulated_data <- rpois(n, lambda = theta_true)
  theta_ML <- mle_poisson(simulated_data)
  results[i] <- sqrt(n) * (theta_ML - theta_true)/sqrt(theta_ML)
}

# Plot histogram
hist(results, main = "Histogram of sqrt(n)(theta_ML - theta)/sqrt(theta_ML)", col = "lightblue", xlab = "Standardized Variable")
abline(v = 0, col = "red", lwd = 2)


```

```{r}
# QQ-Plot against a standardized normal distribution
qqnorm(results)
qqline(results, col = "red", lwd = 2)
```

```{r}
# Optional: Conduct a Shapiro-Wilk test for normality
shapiro.test(results)

```

Since the p-value is greater than the commonly chosen significance level of 0.05, you would not reject the null hypothesis. This means that, based on the Shapiro-Wilk test, there is no strong evidence to suggest that the data significantly deviates from a normal distribution.


The p-value obtained in this analysis exceeds the commonly accepted threshold of 0.05, a standard significance level in statistical tests. Consequently, this leads us to retain the null hypothesis. The implication of not rejecting the null hypothesis, particularly in the context of the Shapiro-Wilk test, is significant. It indicates that the evidence is not strong enough to conclude that the data deviate significantly from a normal distribution.

In simpler terms, as the p-value is higher than 0.05, it suggests that the observed data are consistent enough with a normal distribution.